---
title: 'STAT 311 Final Project Part 1: EDA'
author: "Timila Kulkarni"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(alr4)
library(car) # for scatter plot function
library (ggpubr) #for ggarrange function
```

```{r read in Housing data}
Housing <- read_csv("Housing_Group31.csv", show_col_types=FALSE)
Housing$comes_furnished <- factor(Housing$comes_furnished)
levels(Housing$comes_furnished) <- c("not furnished", "furnished")
Housing$cats_allowed <- factor(Housing$cats_allowed)
levels(Housing$cats_allowed) <- c("cats not allowed", "cats allowed")
```

## Categorical Variables

### Univariate explorations

```{r tables of counts as percentages}
# comes_furnished
Housing %>% 
  count(comes_furnished) %>%
  mutate(Prop = round(n / sum(n) * 100, 1)) %>%
  pivot_wider(id_cols = c(comes_furnished, Prop), names_from = comes_furnished, values_from = Prop)

# cats_allowed
Housing %>% 
  count(cats_allowed) %>%
  mutate(Prop = round(n / sum(n) * 100, 1)) %>%
  pivot_wider(id_cols = c(cats_allowed, Prop), names_from = cats_allowed, values_from = Prop)
```

### Bivariate Explorations

#### 2-way contingency tables

```{r 2-way contingency tables}
# with joint counts
Housing %>%
  count(comes_furnished, cats_allowed) %>%
  pivot_wider(names_from = cats_allowed, values_from = n)

# with row conditional percentages
Housing %>%
  count(comes_furnished, cats_allowed) %>%
  group_by(comes_furnished) %>%
  mutate(pct = round(n / sum(n) * 100, 1)) %>% 
  pivot_wider(id_cols = c(comes_furnished, cats_allowed), names_from = cats_allowed, values_from = pct)
```

#### Bar Graphs

```{r Bar Graphs}
#jpeg(file="Bar_Graph.jpg", width = 450, height = 650) 
ggplot(Housing, aes(x = comes_furnished, fill = cats_allowed)) + 
  geom_bar(position = "fill") +
  labs(y = "Proportion")
#dev.off()
```

## Quantitative Variables

### Univariate Explorations

#### 7-number summaries

```{r 7-number summaries}
# price
(sum.price <- Housing %>% summarize(Min = min(price, na.rm=TRUE), 
                                  Q1 = quantile(price, 0.25, na.rm=TRUE),
                                  Median = quantile(price, 0.50,
                                                    na.rm=TRUE),
                                  Q3 = quantile(price, 0.75, na.rm=TRUE),
                                  Max = max(price, na.rm=TRUE),
                                  mean = mean(price, na.rm=TRUE),
                                  SD = sd(price, na.rm=TRUE)))

# number of beds
(sum.beds <- Housing %>% summarize(Min = min(beds, na.rm=TRUE), 
                                  Q1 = quantile(beds, 0.25, na.rm=TRUE),
                                  Median = quantile(beds, 0.50,
                                                    na.rm=TRUE),
                                  Q3 = quantile(beds, 0.75, na.rm=TRUE),
                                  Max = max(beds, na.rm=TRUE),
                                  mean = mean(beds, na.rm=TRUE),
                                  SD = sd(beds, na.rm=TRUE)))
```


#### clean the data
```{r clean up}
Q.price <- quantile(Housing$price, probs=c(.25, .75), na.rm = FALSE)
iqr.price <- IQR(Housing$price)

Q.beds <- quantile(Housing$beds, probs=c(.25, .75), na.rm = FALSE)
iqr.beds <- IQR(Housing$beds)

Q.lat <- quantile(Housing$lat, probs=c(.25, .75), na.rm = TRUE)
iqr.lat <- IQR(Housing$lat, na.rm = TRUE)

new_Housing <- filter(Housing, price > (Q.price[1] - 1.5*iqr.price) & Housing$price < (Q.price[2]+1.5*iqr.price) & beds != 1100 & beds != 1000 & lat > (Q.lat[1] - 1.5*iqr.lat) & Housing$lat < (Q.lat[2]+1.5*iqr.lat))

```

#### Histograms

```{r Histograms}
#price
#jpeg(file="Price_Hist.jpg", width = 450, height = 450) 
ggplot(data = new_Housing, aes(x = price)) +
  geom_histogram(aes(y = ..density..), bins=75, colour = 1, fill = "white") +
  xlab("Price ($ per month)")
#dev.off()

# beds
#jpeg(file="Beds_Hist.jpg", width = 450, height = 450) 
ggplot(data = new_Housing, aes(x = beds)) +
  geom_histogram(colour = 1, fill = "white", bins=10) + 
  xlab("Number of beds")
#dev.off()
```

#### Box Plots

```{r Box Plots}
price_box <- ggplot(Housing, aes(x=log(price), y="")) +
        geom_boxplot() + 
        xlab("Price ($ per month)") +
        ylab("")
price_box

beds_box <- ggplot(Housing, aes(x=beds, y="")) +
        geom_boxplot() + 
        xlab("Number of beds") +
        ylab("")
beds_box
```


### Bivariate Explorations

#### Scatterplot

```{r Scatterplot}
#jpeg(file="Scatterplot.jpg", width = 650, height = 650) 
ggplot(new_Housing, aes(x=beds, y=price)) +
  geom_point() + geom_abline(slope = 1, intercept = 0)
#dev.off()

#jpeg(file="Scatterplot_lat.jpg", width = 650, height = 650) 
ggplot(new_Housing, aes(x=lat, y=price)) +
  geom_point() + geom_abline(slope = 1, intercept = 0)
#dev.off()
```


## Quantitative and Categorical Variables

#### Faceted Histogram

```{r Faceted Histogram}
# Price by Cats Allowed
ggplot(data = new_Housing, aes(x = price)) +
  geom_histogram(aes(y = ..density..), bins = 50, colour = 1, fill = "white") + 
  xlab("Price ($ per month)") + facet_wrap(vars(cats_allowed))

# Price by Comes Furnished
#jpeg(file="Furnished_Hist.jpg", width = 650, height = 650) 
ggplot(data = new_Housing, aes(x = price)) +
  geom_histogram(aes(y = ..density..), bins = 50, colour = 1, fill = "white") + 
  xlab("Price ($ per month)") + facet_wrap(vars(comes_furnished))

#dev.off()


# Beds by Comes Furnished
ggplot(data = new_Housing, aes(x = beds)) +
  geom_histogram(aes(y = ..density..), bins = 10, colour = 1, fill = "white") + 
  xlab("Number of Beds") + facet_wrap(vars(comes_furnished))

#Beds by Cats Allowed
ggplot(data = new_Housing, aes(x = beds)) +
  geom_histogram(aes(y = ..density..), bins = 10, colour = 1, fill = "white") + 
  xlab("Number of Beds") + facet_wrap(vars(cats_allowed))
```

#### Comparative Boxplots

```{r Comparative Boxplots}
#jpeg(file="BoxPlot.jpg", width = 700, height = 650) 
ggplot(new_Housing, aes(x=comes_furnished, y=price)) + 
  geom_boxplot() + xlab("Comes Furnished") +
  ylab("Price ($ per month)")
#dev.off()